---
title: "Loop for simulation"
author: "Charlotte Brault"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
colorlinks: true
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 5
    number_sections: TRUE
urlcolor: blue
---
```{r setup, include=FALSE}
R.v.maj <- as.numeric(R.version$major)
R.v.min.1 <- as.numeric(strsplit(R.version$minor, "\\.")[[1]][1])
if(R.v.maj < 2 || (R.v.maj == 2 && R.v.min.1 < 15))
  stop("requires R >= 2.15", call.=FALSE)

suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE, fig.align="center")
```


# Overview

Importantly, before anything else, one must specify all paths relatively to the root of the project:
```{r project_organization}
project.name <- "mv-sxg-pharch"
project.dir <- "~/mv-sxg-pharch"
if(Sys.info()["user"] == "tflutre"){
  project.dir <- paste0("~/work/", project.name)
} else if(Sys.info()["user"] == "flutre"){
  project.dir <- paste0("/work/", project.name)
} else if(Sys.info()["user"] == "jchiquet"){
  project.dir <- paste0("~/git/", project.name)
} else if(Sys.info()["user"] == "braultch"){
  project.dir <- paste0("C:/Users/braultch/work/", project.name)
} else if(Sys.info()["user"] == "brault"){
  project.dir <- paste0("~/work/", project.name)
} else if(Sys.info()["user"] == "perrot-dockes"){
  project.dir <- paste0("/home/perrot-dockes/Documents/Multivar_selec/Multivar_selec/FusedLasso/Charlotte/", project.name)
}
stopifnot(file.exists(project.dir))
data.dir <- paste0(project.dir, "/data")
stopifnot(file.exists(data.dir))
results.dir <- paste0(project.dir, "/results")
stopifnot(file.exists(results.dir))
```

This document will also require external packages to be available, for instance:
```{r load_pkg}
suppressPackageStartupMessages(library(beanplot))
suppressPackageStartupMessages(library(qtl))
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(pbmcapply))
suppressPackageStartupMessages(library(glmnet))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(stabs))
suppressPackageStartupMessages(library(Fus2mod))
suppressPackageStartupMessages(library(lars))
suppressPackageStartupMessages(library(spring))
nb.cores <- detectCores() - 1
if(Sys.info()["sysname"] == "Windows")
  nb.cores <- 1
if(nb.cores > 20){
  nb.cores <- 20
}
suppressPackageStartupMessages(library(rutilstimflutre))
source(file = paste0(project.dir,"/src/Function4GenoClass.R"))
 source(paste0(project.dir,'/src/grp_lasso.R'))
```

This R chunk is used to assess how much time it takes to execute the R code in this document until the end:
```{r time_0}
t0 <- proc.time()
```

We are trying to define which method is the best for multivariate QTL detection based in simulated data.
For more details, see the Rmd `mv-sxg-pharch_comparison-QTL-detection`



# Marker genotypes

## As genotypic class

From Doligez et al (2013), these data are used to perform QTL analysis through simple tests.
Contains SSR genotypes (`.loc`), genetic map (`.map`) and some quantitative phenotypes (`qua`, not used in this project) in the [JoinMap](https://www.kyazma.nl/index.php/mc.JoinMap/) format, readable by the `qtl` package.

This file is used for the univariate QTL detection.

```{r read_cross}
dat.doligez2013 <- read.cross(format = "mapqtl", dir=data.dir,
                              genfile="SxG_Doligez-et-al-2013.loc",
                              mapfile = "SxG_Doligez-et-al-2013.map",
                              phefile="SxG_Doligez-et-al-2013.qua")
```

Get genotypic classes of all markers as a data frame:

```{r results="hide"}
p2f <- paste0(data.dir, "/SxG_Doligez-et-al-2013.loc")
dim(gclasses <- read.table(p2f, skip=5))
rownames(gclasses) <- gclasses[,1]
gclasses <- gclasses[,-1]
colnames(gclasses) <- c("seg", "phase", as.character(dat.doligez2013$pheno$indiv))

tmp = data.frame(locus=rownames(gclasses),
                 seg=gclasses$seg,
                 phase=gclasses$phase,
                 clas=NA)
tmp = cbind(tmp, gclasses[, -c(1:2)])
tmp = updateJoinMap(tmp)
gclasses = tmp[, -c(1,4)]
clgeno = gclasses
```


Construct the genetic map, for each chromosome list the marker and its position in gmap (list)

```{r read_genetic_map}
map <- read.table(paste0(data.dir,
                         "/SxG_Doligez-et-al-2013.map"),
                  stringsAsFactors=FALSE)
tmp <- map
link.groups <- as.character(tmp[,2][tmp[,1] == "group"])
gmap <- list()
for(link.group in link.groups[1:(length(link.groups) - 1)]){
  idx.start <- which(tmp[,2] == link.group) + 1
  idx.end <- which(tmp[,1] == "group" &
                   tmp[,2] != link.group)[1] - 1
  gmap[[link.group]] <- setNames(as.numeric(tmp[idx.start:idx.end, 2]),
                                 tmp[idx.start:idx.end, 1])
  tmp <- tmp[-((idx.start-1):idx.end),]
}
link.group <- tail(link.groups, 1)
gmap[[link.group]] <- setNames(as.numeric(tmp[2:nrow(tmp), 2]),
                               tmp[2:nrow(tmp), 1])
```


Get the linkage groups from the marker names, thus it is possible to identify the linkage group of each marker.

```{r get_lg}
mrk.names <- do.call(c, lapply(gmap, names))
rep.lg.names <- do.call(c, lapply(names(gmap), function(lg.name){
  rep(lg.name, length(gmap[[lg.name]]))
}))
mrk2lg <- setNames(rep.lg.names, mrk.names)
```


## Design matrix of marker genotypes

For this, `mv-sxg-pharch_reformat-genos.Rmd` must have been executed beforehand.

```{r read_X_genos}
X <- as.matrix(read.table(paste0(results.dir,
                                 "/SxG_Doligez2013_genos-design-mat.tsv"),sep="\t"))
X <- X[,-1] #remove intercept
dim(X)
X[1:3, 1:5]
stopifnot(! any(is.na(X)))
```



## Genetic map to data frame

The function `spring` is taking as genetic map a data frame with a column chrom containing the chromosome, a column loci with genetic distance and names with marker names.

We have to transform the file gmap to a data frame in the right configuration.


```{r gmap2df}
gmapdf <- data.frame(chrom=0, loci=0, names=mrk.names)
for (i in 1:nrow(gmapdf)){
  gmapdf$chrom[i] <- substr(rownames(gmapdf)[i], start=2, stop=3)
  idx <- which(map$V1 == gmapdf$names[i])
  gmapdf$loci[i] = map$V2[idx]
}
rownames(gmapdf) <- NULL
```

# Load phenotypes

Phenotypes are coming from the work of Aude Coupel Ledru, for more details, see her article in 2014 and 2016.


Traits are taken under 2 conditions and 2 years 

In Coupel Ledru et al, there is a QTL for the trait KS in years 12+13 for the WW scenario in the linkage group 2 near the marker VVIB01

```{r load_pheno}
blups = read.table(file=paste0(results.dir,"/SxG_CoupelLedru2014_all-blups.tsv"), header=TRUE, sep="\t", dec='.')
head(blups)
trait <- "KS"
year <- "12+13"
scenario <- c("WW","WD")

# here year is fixed so the multivariate aspect is coming from the condition
Condition1 = blups[which(blups$trait == trait & blups$year == year & 
                  blups$scenario == scenario[1]),"blup"]

Condition2 = blups[which(blups$trait == trait & blups$year == year & 
                  blups$scenario == scenario[2]),"blup"]

Y = cbind(Condition1,Condition2)

rownames(Y) <- blups[which(blups$trait == trait & 
                              blups$year == year & 
                              blups$scenario == scenario[1]),"geno"]
head(Y)

marker <- "VVIB01"
```


# Marker selection

```{r loop}
res_G <- 0
res_Fus <- 0 
res_Spring <- 0
U1 <- data.frame(chromosome=0, position=0, LOD=0, idx=0) 
U2 <- data.frame(chromosome=0, position=0, LOD=0, idx=0)

### Plot data
setd1 = c(setdiff(colnames(clgeno),rownames(Y)))
setd2 = c(setdiff(rownames(Y),colnames(clgeno)))
setd3 = c(setdiff(rownames(Y),rownames(X)))
setd4 = c(setdiff(rownames(X),rownames(Y)))

clgeno = clgeno[, !(names(clgeno) %in% setd1)]
Y = Y[ ! (rownames(Y) %in% setd2),]
Y = Y[ ! (rownames(Y) %in% setd3),]
X = X[ ! (rownames(X) %in% setd4),]
(dim(X))
(dim(Y))

clgeno[clgeno == '--'] <- NA
clgeno.marker = as.matrix(clgeno[rownames(clgeno) == marker,])
ylim = c(-2,2)

par(mfrow = c(1,2))
 
beanplot(Y[,"Condition1"] ~ t(clgeno.marker), ylim=ylim, xlab = "genotypic class", 
       ylab="Simulated response", main = "Response 1",las=1, log="", bw="nrd0", 
       what=c(1,1,1,0), col=c("grey",0,0,"red"), border=NA)

beanplot(Y[,"Condition2"] ~ t(clgeno.marker), ylim=ylim, xlab="genotypic class", 
       ylab="Simulated response", main="Response 2",las=1, log="", bw="nrd0", 
       what=c(1,1,1,0), col=c("grey",0,0,"red"), border=NA)

par(mfrow = c(1,1))

## Univariate QTL detection
  ### Response 1
U.resp1 = univQTL(dat.doligez2013, response=subset(Y,select="Condition1"), 
              nperm=100, method ="ehk", plot=TRUE, ylim=c(0,15),
             main = paste0(trait,"-", year, " - WW"))
if(nrow(U.resp1) == 0) {
    U1 = 0
    
  } else if (nrow(U.resp1) != 0) {
  U1 = U.resp1
  }
  
### Response 2
U.resp2=univQTL(dat.doligez2013, response=subset(Y,select="Condition2"),
          nperm=100, method ="ehk", plot=TRUE, ylim=c(0,15), 
           main = paste0(trait,"-", year, " - WD"))

if(nrow(U.resp2) == 0){
    U2 = 0

    
    }else if(nrow(U.resp2) != 0){
    U2 = U.resp2
}
U1
U2

## Glmnet
G = glmnetmultivarQTL(X=X, Y=Y, map=map, choose.alpha=0.1, 
                  PFER=1, B=500, cutoff=0.95, plot=TRUE, nb.cores=nb.cores)
res_G = G

## Fus2mod
# group = c("WW","WD")
# Y <- as.data.frame(Y)
# Fu=fus2modQTL(X=X,Y=Y, map=map, group=group, 
#               nrep=c(2,10), plot=c(FALSE,TRUE), cutoff=0.95, lambda=NULL, nb.cores=nb.cores)
#   
# res_Fus = Fu[[1]]
# (print(Fu[[2]]))
#   
## Spring
Sp = springQTL(X=X, Y=Y, map=map, gmapdf=gmapdf, group=group,
                 plot=c(FALSE),lambda2=c(10^seq(2,-2,len=6)),nb.cores)
res_Spring <- Sp

Sp
## Group lasso
Glb <- grpLassoQTL(X, Y, marker, map,
                        PFER=1, B=500, cutoff=0.85, plot=TRUE, nb.cores, sep ="\\.",  mrk2lg, type_group="both")
   
res_Glb<- rbind(res_Glb, Glb)
Glb$idx <- i

Glb
  
    Glt <- grpLassoQTL(X, Y, marker, map,
                        PFER=1, B=500, cutoff=0.85, plot=TRUE, nb.cores, sep ="\\.", mrk2lg, type_group="Trait")

Glt$idx <- i
      res_Glt <- rbind(res_Glt, Glt)

        Glt

```


# Appendix


```{r info}
format(Sys.time(), '%d/%m/%Y %H:%M:%S')
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```






